---
title: "Scenario Management Vignette"
output:
  html_document: 
    theme: cosmo
---

*scenario_management.R (defines ve.scenario_management tools)*

This file documents the `ve.scenario_management` tools for VisionEval, which allows users to manage scenarios using a CSV file rather than JSON files.  

## Background

A powerful feature of VisionEval is the ability to quickly run 100s or even 1000s of scenarios. Background on running scenarios in VisionEval is available in the [project documentation](https://github.com/VisionEval/VisionEval/wiki/Multiple-Scenarios). Scenarios are defined using a `JSON` file named `scenario_config.json`. In your distribution of VisionEval, you can see an example of this file in the `sources/models/VERPAT_Scenarios/scenario_inputs/` directory. A second `JSON` file, `category_config.json`, defines which of the scenarios a user wants to run. Having this second file allows the user to easily run a subset of the universe of defined scenarios. 

## Scenario Management Tools Demonstration
`JSON` files are easily consumed by software, but difficult to edit manually. The scenario management tools provides an alternative to editing the JSON file in the form of a database. 

Here's how the tools work.

To access the tools you must, let R know where they are. The code below sets the working directory to `VisionEval/sources/`. Please modify the code to point to where you have saved the `VisionEval/sources` folder.
```{r setup}
knitr::opts_knit$set(root.dir = "~/Documents/GitHub/VisionEval-Dev/sources/")
```

We can now load the VisionEval scenario management tools.
```{r load-tools, message = FALSE}
source("tools/scenario_management.R")
```

Let's start with the example VERPAT Scenario configuration files that are distributed with VisionEval. The code below tells the software where these files are.   
```{r example-file-dir}
example_file_dir <- "models/VERPAT_Scenarios/scenario_inputs/"
```

We can build a Standard VisionEval scenario management CSV database from the `JSON` files located in this directory using the example files distributed with VisionEval. You can name the CSV database anything you like and store it any place you like.
```{r example-csv-build}
example_csv_database_file <- "tools/example_scenario_manage.csv"
ve.scenario_management.make_form_csv_from_json(example_file_dir, example_csv_database_file)
```

Let's bring the newly-created CSV file into R to see what it looks like.
```{r view-form}
example_form_df <- read_csv(example_csv_database_file)
```

This CSV file can now be edited in Excel, a text editor, R, or any other software. The `scenario_management` tools allow you to build the `JSON` files that VisionEval needs from these CSV files. For example, let's first specify a new directory location for the `JSON` files we'll make and then use the tools to build them:
```{r build-json}
new_json_file_dir <- "tools/example_dir/"
ve.scenario_management.make_json_from_form_csv(example_csv_database_file, new_json_file_dir)
```

The new `JSON` files can be viewed in a text editor or in R Studio, among other places. We can use R to test that the `JSON` files we've made are the same -- when read by a computer -- as the `JSON` files we started with using R's `identical` method, like this:
```{r check-identical}
identical(read_json(file.path(example_file_dir, "scenario_config.json")), 
          read_json(file.path(new_json_file_dir, "scenario_config.json")))

identical(read_json(file.path(example_file_dir, "category_config.json")), 
          read_json(file.path(new_json_file_dir, "category_config.json")))
```

Another tedious requirement of the `VEScenario` package is that it requires input files to be stored in a specific, nested directory structure. Making this directory structure by hand is tedious and error-prone. The scenario management tools now do this for you, like this:
```{r make-dir}
new_scenario_folders_dir <- "tools/example_dir/"
ve.scenario_management.make_directory_structure(new_scenario_folders_dir, example_csv_database_file)
```



