---
title: "Introduction to VERRSPM_Scenario"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to VETRSPM_Scenario}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

Strategic planning often requires the assessment of large numbers of future scenarios, each assessing a different combination of prioritizations, policy decisions, and constraints. The VEScenarios module provides the capability to quickly set up, run, and visualize large numbers of VERSPM scenarios using a baseline scenario combined with multiple changes to model inputs. 

Both VERSPM and VERPAT models are capable of running multiple scenarios efficiently


### Scenario manager database
The first step to run the VERSPM_scnerios is to build the scenario manager database. 
The database should be csv file in standard defined in `defs/verspm_scenario_manager.csv`. You can 
add as many sceanrios as you want but the format of each row should remain the same.

A sample file is in `defs/verspm_scenario_manager.csv`. 
At first level, you should define the strategies you want to run. 

These are the default list of VERSPM_Scenarios strategies:


* **Community Design** :  Policies that seek to enable shorter trips and alternate modes such as promotion of mixed use land use, transit service, bicycling, and parking management. ( group of L, B, P , T ) 
* **Marketing/Incentive** : Policies that improve driving efficiency such as eco driving, and Intelligent Transportation System efforts, as well as programs that reduce auto demand such as carsharing, and home or work-based transportation demand management.( group of D, E ) 
* **Pricing : Policies** that move towards true cost pricing such as road user fees to pay for the cost of operating, maintaining and improving roads, pay-as-you-drive (PAYD) insurance, and environmental impact fees such as a carbon tax. ( C )
* **Vehicles/Fuels** : Factors representing changes to future vehicles and fuels ( group of V, F ) 
* **Income** : Context factor on the assumed growth of statewide average per capita income ( I )

Each strategy covers couple of categories in a lower level.

These are the current list of VERSPM_Scenarios categories:

  - B - Bicycles (`azone_prop_sov_dvmt_diverted.csv`) : Network improvements, incentives, and technologies that encourage bicycling and other light-weight vehicle travel

  - C - Vehicle Travel Cost (`azone_hh_veh_own_taxes.csv`) and (`region_prop_externalities_paid.csv`) : Combination of fuel prices and charges to pay for roadway costs and possibly externalities

  - D - DemandManagement (`bzone_travel_demand_mgt.csv`) : Programs to encourage less private vehicle travel

  - E - Driving Efficiency (`marea_operations_deployment.csv`) and (`marea_speed_smooth_ecodrive.csv`) and (`other_ops_effectiveness.csv`) : Driving efficiency by increasing implementation of ITS

  - F - Technology Mix and CI (`marea_transit_powertrain_prop.csv`) and (`region_carsvc_powertrain_prop.csv`) and (region_comsvc_powertrain_prop.csv`) : Vehicle technology mix and carbon intensity of fuels.

  - G - Fuel Price (`azone_fuel_power_cost.csv`) : Real fuel price in 2010 USD

  - I - Income (`azone_fuel_power_cost.csv`) : Real average household income in 2010 USD

  - L - LandUse (`azone_hhsize_targets.csv`) and (`bzone_urban_du_proportions.csv`) : Distribution of population and employment by place type

  - P - Parking (`bzone_parking.csv`) : The extent of paid parking and its price

  - T - Transit (`marea_transit_service.csv`) : The extent and frequency of transit service

  - V - Vehicle Characteristics (`azone_hh_veh_mean_age.csv`) and (`azone_lttrk_prop`): The combination of fuel prices and vehicle travel charges to pay for roadways and to pay for externalities such as carbon pricing

 For each category, you should define the levels you would like to run. For each category level you need to change a set of inputs based on the scenario definition. 
 
 The input column should be populated with the input file names for each scenario level.

You can compeletly delete the default list and add new strategies for your own specific scenarios.

After creating this database when you run the VERSPM_Scenarios model, a directory with all the scneario levels and required inputs will be created . see `scenario_inputs` for an example of the directory structure.

### defs

`VERSPM_Scenarios/defs` directory contains the same files as the `/defs` directory in VERSPM, but the `model_parameters.json` is different.:

```
[
  

  {
    "NAME": "ModelFolder",
    "VALUE": "VERSPM_base_model",
    "TYPE": "character",
    "UNITS": "NA",
    "PROHIBIT": "NA",
    "SIZE": 20,
    "ISELEMENTOF": ""
  },
  {
    "NAME": "ScenarioInputFolder",
    "VALUE": "scenario_inputs",
    "TYPE": "character",
    "UNITS": "NA",
    "PROHIBIT": "NA",
    "SIZE": 20,
    "ISELEMENTOF": ""
  },
  {
    "NAME" : "ScenarioManagerFile",
    "VALUE": "verspm_scenario_manager.csv",
   "TYPE": "character",
   "UNITS": "NA",
   "PROHIBIT": "NA",
   "SIZE": 30,
   "ISELEMENTOF": ""
  },
  {
    "NAME" : "ScenarioOutputFolder",
    "VALUE": "scenarios",
   "TYPE": "character",
   "UNITS": "NA",
   "PROHIBIT": "NA",
   "SIZE": 20,
   "ISELEMENTOF": ""
  },  
  {
    "NAME" : "NWorkers",
    "VALUE": "6",
    "TYPE" : "integer",
    "UNITS" : "NA",
    "PROHIBIT" : "c('NA', '< 0')",
    "ISELEMENTOF" : ""
  }
]

                                                                                                                                                                                                                                                         
```
** Model folder : The VERSPM base model folder name 
** ScenarioInputFolder: Folder name for scenario root where scenario manager file is located. All the scenario level folders will be created here
** ScenarioManagerFile : Scenario manager database file name
** ScenarioOutputFolder: Name of the folder where all the scenarios are created


### Modules
Running VERSPM_Scenario has 5 major modules:



** NewScenarioSet : This module creates the required categories and scenarios json files from csv database. It also creates the folder structure and required inputs for all scenario levels  
** BuildScenarios: This module creates all possible scnearios which are defined in previous step. It copies the model files from the base VERSPM model folder for each scenario.
** RunScenarios : This module runs all scenarios 
** VERSPMResults: This model summarizes all different scenario outputs from datastore to standard format
** ViewResults:  This module creates the visualization and summaries for different performance measures 



### Running VERSPM_Scenarios

After opening `VisionEval-dev` project file, follow the following steps to run the VERSPM_Scenarios model. The model run has two stages:

Stage 1: Creates scenario_inputs folder based on the scenario manager database and places all the UNEDITED model inputs into sub-directories for all scenario levels. Run the following scripts in console for stage 1:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ve.run()
sample_model_run <- openModel("VERSPM_Scenarios")
sample_model_run$run(stage = 1)

```

After you see the "stage-1 completed" message, you should replace the model Unedited versions with your revised data. go to the scenario inputs folder and for each category level find the unedited input belong to each level. Modify the file with your specified inputs and delete "unedited" part from the file names. After you update all the inputs, you can run the following scripts in console:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
sample_model_run$run(stage = 2)

```


